---
title: "CH 4.3"
author: "Patrick Wong"
date: "10/31/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MATH 324 Ch 4.2

##Standard Normal R Calculations

##1A) P(0 < Z < 2.17)?
```{r}
pnorm(2.17) - pnorm(0)
```

##1B) P(-2.5 <= Z <= 0)?
```{r}
pnorm(0) - pnorm(-2.5)
```

##1C) P(-2.5 <= Z <= 2.5)?
```{r}
pnorm(2.5)-pnorm(-2.5)
```
##1D) P(Z <= 1.37)?
```{r}
pnorm(1.37)
```

##1E) P(Z >= -1.75)?
```{r}
1-pnorm(-1.75)
```

##1G) P(-1.5 <= Z <= 2)?
```{r}
pnorm(2)-pnorm(-1.5)
```

##1H) P(1.37 <= Z <= 2.5)?
```{r}
pnorm(2.5)-pnorm(1.37)
```

##1I) P(Z >= 1.5)?
```{r}
1-pnorm(1.5)
```


##1J) P(|Z| <= 2.50) = P(-2.50 <= Z <= 2.50)
```{r}
pnorm(2.5)-pnorm(-2.5)
```

## Percentiles of Z

##2A) n(91)
phi(z) = 0.91 = c = 1.34 

##2B) n(9)
phi(z) = 0.09 = -0.91 = -1.34

##2C) n(75)
phi(z) = 0.75 = c =.675

##2D) n(25)
phi(z) = 0.25 = -0.75 = -0.675

##2E) n(6)
phi(z) = 0.06 = c = -0.1555

##3A) phi(c) = 0.9838 (98.38th percentile)
found in 2.1 row and the 0.04 column so standard table = c = 2.14

##3B) P(0 <= Z <= c)= 0.291
phi(c) = 0.7910 = c =0.81

##3C) 1 - phi(c)=P(Z >= c)=.121
 1 - P(c <= Z) = P(Z < c) = phi(c)= 1 - 0.121 = 0.8790 = c = 1.17
 
##3D) P(-c <= Z <= c)= 0.668
P(-c<= Z <= c)= phi(c)- pho(-c) = phi(c) – (1 - phi(c)) = 2 * phi(c) – 1 = 
phi(c)= 0.9920 = c = 0.97

##3E) P(|Z| >= c)=.016
1 - 0.016 = 0.9840 = 1–P(c <= |Z|) = P(|Z| < c) = P(-c < Z < c) = 
phi(c) - phi(-c) = 2 * phi(c) – 1 = 
phi(c) = 0.9920 = c = 2.41

##4A) alpha = 0.0055 (99.45th percentile)
```{r}
qnorm(0.055)
```

##4B) alpha = 0.09 (100(1-.09)th percentile)
```{r}
qnorm(0.09)
```

##4C) alpha = 0.663 (100(1-.66)th percentile)
```{r}
qnorm(0.663)
```

##Word Problems for X ~ N(Mu, sigma ^ 2)

In a road-paving process, asphalt mix is transfered by truck. 
Let X = truck haul time be normally distributed with mean 8.46 min 
and standard deviation .913 min.

##5A)
What is the probability that haul time will be at least 10 minutes?

P(X >= 10) = P((X - Mu/sigma) >= 10-8.46/0.913)
P(Z >= 1.69)
P(Z <= -1.69)
```{r}
pnorm(-1.69)
```

##5B)
What is the probability that haul time will exceed 10 minutes?

P(X > 10) = P(X >= 10)
= 0.0455

##5C) 
What is the probability that haul time will be between 8 and 10 minutes?

P(8 <= X <= 10) = P((8-8.46/0.913) <= (X - Mu/sigma) <= (10-8.46/0.913))
= P(-0.5 <= Z <= 1.69)
= P(Z <= 1.69) - P(Z <= -0.5)
```{r}
pnorm(1.69) - pnorm(-0.5)
```

##5D)
What value c is such that 98% of all haul times are in the interval 
from 8.46 - c to 8.46 + c.

P(8.46 - c <= X <= 8.46 + c) = 0.98
P(-c <= X -8.46 <= c) = 0.98
P(-c <= X - Mu <= c) = 0.98
P((-c/0.913) <= (X - Mu/sigma) <= (c/0.9.13) = 0.98
P((-c/0.913 <= Z <= (c/0.913)) = 0.98
2P (Z <= c/0.913) - 1 = 0.98
P(Z <= c/0.913) = 1+0.98/2
P(Z <= c/0.913) = 0.99
c/0.913 = 2.33
c = 2.33 * 0.913
c = 2.127


##5E)
If four haul times are independently selected, what is the probability that at least one of them
exceeds 10 minutes?

P(X > 10) = 0.0455
P(at least on exceeds) = 1 - P(none of the four exceeds)
= 1 - P(X<10)^4
= 1 - [1 - P(X > 10)]^4
= 1 - [1 - 0.0455]^4
= 0.169951
